---
title: "Problem Sheet 2"
author: "Your Name (Reg No)"
subtitle: Mean Vectors & Covariance Matrices
output:
  html_notebook: default
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, echo=FALSE}
library(lemon) #This is to show the table pretty in pdf
knit_print.data.frame <- lemon_print
```

Consider the data set `crossSecData`(You may download the data set fom [https://github.com/davidrajdgl/MultivariteAnalysis/blob/master/DataSets/crossSecData.xlsx](https://github.com/davidrajdgl/MultivariteAnalysis/blob/master/DataSets/crossSecData.xlsx). The details of the variables are given below. 

- Id : Patient Id
- Diabetes: Diabetes Present : 1; Obsent : 0
- Fat_intake: The amount of fat intake
- Height : Height of the patient
- Weight : Weight of the patient
- SBP : Systollic Blood Pressure
- DBP : Diastolic blood pressure

**Important:** Dont forget to save the dataset in your working directory. Set the working directory before you proceed to answer any questions below. 


**Load the data set Here:**
```{r, caption ="First 5 rows of the Data Set"}
library(readxl)
data= read_excel('crossSecData.xlsx')
head(data, n = 5) #Show only first 5 rows of the data
```

**Question 1** What is a Random vector?




**Question 2:** What is the Random vector in this data set?

*Answer:*

```{r, message=FALSE, error=FALSE}
randVec = colnames(data)
randVec
```

The random vector of the data is  (`r randVec` )$^T$.

**Question 3: **  Generate a new data set from the `data`, consists of Fat_intake, Height, SBP and DBP only. Save the data set as 'newCrossSec' and print only first 10 rows of the new data. 

*Answer:*  

```{r}
newCrossSec = data[ , c(3, 4,6,7)]

head(newCrossSec, n = 10)
```


**Question 4: ** What is the Random Vector in the new data set?

*Answer:*  




**Question 5:**  What is the sample mean vector in the new data set? 

*Answer:*  
```{r}
#Load the new data set as X 
X  =  newCrossSec
#using the matrix approach find the mean vector. Note that you have to  generate a vector of all ones. Hint: vector of all 2's can be acheived by `rep(2, 100)`
one = rep(1, nrow(X))
SampleMeanVec = 1 / nrow(X) * t(X) %*% one
SampleMeanVec

```


#### Some useful hints: 

Remember to find the sample covariance matrix, we need to find the matrxi $X^*$ where 


\begin{equation}
X^* = \begin{pmatrix}
X_{11} - \overline{X_1} & X_{12} - \overline{X_2} & \dots & X_{1p} - \overline{X_p} \\
X_{21} - \overline{X_1} & X_{22} - \overline{X_2} & \dots & X_{2p} - \overline{X_p} \\
\vdots                  & \vdots                  & \ddots& \vdots                  \\
X_{n1} - \overline{X_1} & X_{n2} - \overline{X_2} & \dots & X_{np} - \overline{X_p} \\
\end{pmatrix}
\end{equation}

Note that, this matrix can be obtained by 

\begin{align}
X^* & = 
\begin{pmatrix}
X_{11} -\overline{X_1} & X_{12} - \overline{X_2} & \dots & X_{1p} - \overline{X_p} \\
X_{21} -\overline{X_1} & X_{22} - \overline{X_2} & \dots & X_{2p} - \overline{X_p} \\
\vdots                & \vdots                  & \ddots& \vdots                  \\
X_{n1} -\overline{X_1} & X_{n2} - \overline{X_2} & \dots & X_{np} - \overline{X_p} \\
\end{pmatrix}\\
& =
\begin{pmatrix}
X_{11} & X_{12} & \dots & X_{1p} \\
X_{21} & X_{22} & \dots & X_{2p} \\
\vdots& \vdots & \ddots & \vdots  \\
X_{n1} & X_{n2}  & \dots & X_{np}  \\
\end{pmatrix} - 
\begin{pmatrix}
\overline{X_1} & \overline{X_2} & \dots & \overline{X_p} \\
\overline{X_1} & \overline{X_2} & \dots & \overline{X_p} \\
\vdots         & \vdots         & \ddots& \vdots         \\
\overline{X_1} & \overline{X_2} & \dots &  \overline{X_p} \\
\end{pmatrix}\\ 
 & =
\begin{pmatrix}
X_{11} & X_{12} & \dots & X_{1p} \\
X_{21} & X_{22} & \dots & X_{2p} \\
\vdots& \vdots & \ddots & \vdots  \\
X_{n1} & X_{n2}  & \dots & X_{np}  \\
\end{pmatrix} - 
\begin{pmatrix}
1 \\ 1 \\ \vdots \\ 1 
\end{pmatrix}_{n\times 1}
\begin{pmatrix} 
\overline{X_1} & \overline{X_2} & \dots & \overline{X_p} 
\end{pmatrix}_{1\times p}
\end{align}




**Question 6:** Find the sample covariance matrix of the new data set step by step using the lecture. 







